%-------------------------------------------------------------------------------

% This file is part of Code_Saturne, a general-purpose CFD tool.
%
% Copyright (C) 1998-2012 EDF S.A.
%
% This program is free software; you can redistribute it and/or modify it under
% the terms of the GNU General Public License as published by the Free Software
% Foundation; either version 2 of the License, or (at your option) any later
% version.
%
% This program is distributed in the hope that it will be useful, but WITHOUT
% ANY WARRANTY; without even the implied warranty of MERCHANTABILITY or FITNESS
% FOR A PARTICULAR PURPOSE.  See the GNU General Public License for more
% details.
%
% You should have received a copy of the GNU General Public License along with
% this program; if not, write to the Free Software Foundation, Inc., 51 Franklin
% Street, Fifth Floor, Boston, MA 02110-1301, USA.

%-------------------------------------------------------------------------------

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Short doc CS class corresponding to article
\documentclass[a4paper,10pt,twoside]{csshortdoc}
% MACROS SUPPLEMENTAIRES
\usepackage{csmacros}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PACKAGES ET COMMANDES POUR LE DOCUMENTS PDF ET LES HYPERLIENS
\hypersetup{%
  pdftitle = {CodeSaturne autovnv},
  pdfauthor = {MFEE},
  pdfpagemode = UseOutlines
}
\pdfinfo{/CreationDate (D:20110704000000-01 00 )}
%
% To have thumbnails upon opening the document under ACROREAD
% pdfpagemode = UseThumbs
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% INFO POUR PAGES DE GARDES
\titreCS{\CS version~\verscs: autovnv tool}

\docassociesCS{}
\resumeCS{This document presents the autovnv tool. The aim of this script
is to drive \CS's cases automatically, to compare checkpoint files and
to display results.

\CS version~\verscs.

\begin{center}
\large{WORK IN PROGRESS}
\end{center}
}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% DEBUT DU DOCUMENT
\begin{document}

\def\contentsname{\textbf{\normalsize TABLE OF CONTENTS}\pdfbookmark[1]{Table of
contents}{contents}}

\pdfbookmark[1]{Flyleaf}{pdg}
\large
\makepdgCS
\normalsize

\passepage

\begin{center}\begin{singlespace}
\tableofcontents
\end{singlespace}\end{center}
%
\section{Introduction}

\textsc{Autovnv} is a small framework to automate the launch of \CS computations
and do some operations on new results.

The script needs a directory of previous \CS cases which are candidates to be
duplicated. This directory is called \textbf{repository}. The duplication is
done in a new directory which is called the \textbf{destination}.

For each duplicated case, \textsc{Autovnv} is able to compile the user files, to
run the case, to compare the obtained checkpoint file with the previous one from
the \textbf{repository}, and to plot curves in order to illustrate the
computations.

For all these steps, \textsc{Autovnv} generate two reports, a global report
which summarizes the status of each case, and a detailed report which gives the
differences between the new results and the previous ones in the
\textbf{repository}, and display the defined plots.

In the \textbf{repository}, previous results of computations are required only
for checkpoint files comparison purpose. They can be also usfull, if the user
needs to run specific scripts.

\section{Installation and prerequisites}

\textsc{Autovnv} does not need a specific installation: the related files are
installed with the other Python scripts of \CS. Additional prerequisites
required are:
\begin{list}{$\bullet$}{}
\item \texttt{numpy},
\item \texttt{matplotlib}.
\end{list}

\section{Command line options}

The command line options can be found with the command: \texttt{code\_saturne
autovnv -h}.

\begin{list}{$\bullet$}{}
\item \texttt{-f FILE, --file=FILE}: gives the file of parameters for
\textsc{Autovnv}. This file is mandatory, and therefore this option must be
completed;
\item \texttt{-q, --quiet}: does not print status messages to stdout;
\item \texttt{-r, --run}: runs all cases;
\item \texttt{-c, --compare}: compares chekpoint files between
\textbf{repository} and \textbf{destination};
\item \texttt{-p, --post}: postprocess results of computations;
\item \texttt{-m ADDRESS1 ADDRESS2 ..., --mail=ADDRESS1 ADDRESS2 ...}: addresses
for sending the reports.
\end{list}

\underline{Examples:}

\begin{list}{$\bullet$}{}
\item \texttt{code\_saturne autovnv -f sample.xml}: duplicates all cases from
the \textbf{repository} in the \textbf{destination}, compile all user files
and exits;
\item \texttt{code\_saturne autovnv -f sample.xml -r}: as above, and run all
cases if defined in \texttt{sample.xml};
\item \texttt{code\_saturne autovnv -f sample.xml -r -c}: as above, and compares
all new checkpoint files with those from the \textbf{repository} if defined
in \texttt{sample.xml};
\item \texttt{code\_saturne autovnv -f sample.xml -r -c -p}: as above, and plots
results if defined in \texttt{sample.xml};
\item \texttt{code\_saturne autovnv -f sample.xml -r -c -p -m
"dt@moulinsart.be dd@moulinsart.be"}: as above, and send the two
reports.
\item \texttt{code\_saturne autovnv -f sample.xml -c -p}: compares and plots
results in the \textbf{destination} already computed.
\end{list}

\underline{Note:}

The detailed report is generated only if the options \texttt{-c, --compare}
or \texttt{-p, --post} is present in the command line.

\section{File of parameters}

The file of parameters is a XML formatted ascii file.

\subsection{Begin and end of the file of parameters}

This example shows the four mandatory first lines of the file of parameters.

\small
\begin{verbatim}
<?xml version="1.0"?>
<autovnv>
    <repository>/home/dupond/codesaturne/MyRepository</repository>
    <destination>/home/dupond/codesaturne/MyDestination</destination>
\end{verbatim}
\normalsize

The third and fourth lines correspond to the definition of the
\textbf{repository} and \textbf{destination} directories.
Inside the markups \texttt{<repository>} and \texttt{<destination>} the user
must inform the related directories. If the \textbf{destination} does not exit,
the directory is created.

The last line of the file of parameters must be:

\small
\begin{verbatim}
</autovnv>
\end{verbatim}
\normalsize

\subsection{Case creation and compilation fo the user files}

When \textsc{Autovnv} is launched, the file of parameters is parsed in order to
known which studies and cases from the \textbf{repository} should be duplicated
in the \textbf{destination}. The selection is done with the markups
\texttt{<study>} and \texttt{<case>} as the following example:

\small
\begin{verbatim}
<?xml version="1.0"?>
<autovnv>
    <repository>/home/dupond/codesaturne/MyRepository</repository>
    <destination>/home/dupond/codesaturne/MyDestination</destination>

    <study label="MyStudy1" status="on">
        <case label="Grid1" status="on" compute="on" post="off"/>
        <case label="Grid2" status="off" compute="on" post="off"/>
    </study>
    <study label="MyStudy2" status="off">
        <case label="k-eps" status="on" compute="on" post="off"/>
        <case label="Rij-eps" status="on" compute="on" post="off"/>
    </study>
</autovnv>
\end{verbatim}
\normalsize

The attributes are:
\begin{list}{$\bullet$}{}
\item \texttt{label}: the name of the file of the script;
\item \texttt{status}: must be equal to \texttt{on} or \texttt{off},
activate or desactivate the markup;
\item \texttt{compute}: must be equal to \texttt{on} or \texttt{off},
activate or desactivate the computation of the case;
\item \texttt{post}: must nb equal to \texttt{on} or \texttt{off},
activate or desactivate the post-processing of the case.
\end{list}

All these attributes are mandatory.

With the attribute \texttt{status}, a single case or a complete study can be
switched off. In the above example, only the case \texttt{Grid1} of the study
\texttt{MyStudy1} is going to be created.

After the creation of the directories in the \textbf{destination}, for each
case, all user files are compiled. The \textsc{Autovnv} stops if a compilation
error occurs: neither computation nor comparison nor plot will be performed,
even if they are switched on.

\underline{Notes:}

\begin{list}{$\bullet$}{}
\item During the duplication, every files are copied, except mesh files, for
which a symbolic link is used.
\item During the duplication, if a file already exists in the
\textbf{destination}, this file is not overwritten by \textsc{Autovnv}.
\end{list}


\subsection{Run cases}

The computations are activated if the option \texttt{-r, --run} is present in
the command line.

All cases described in the file of parameters with the attribute
\texttt{compute="on"} are taken into account.

\small
\begin{verbatim}
<?xml version="1.0"?>
<autovnv>
    <repository>/home/dupond/codesaturne/MyRepository</repository>
    <destination>/home/dupond/codesaturne/MyDestination</destination>

    <study label="MyStudy1" status="on">
        <case label="Grid1" status="on" compute="on" post="off"/>
        <case label="Grid2" status="on" compute="off" post="off"/>
    </study>
    <study label="MyStudy2" status="on">
        <case label="k-eps" status="on" compute="on" post="off"/>
        <case label="Rij-eps" status="on" compute="on" post="off"/>
    </study>
</autovnv>
\end{verbatim}
\normalsize

After the computation, if no error occurs, the attribute \texttt{compute} is set
to \texttt{"off"} in the copy of the file of parameters in the
\textbf{destination}. It is allow to restart \textsc{Autovnv} without re-run
successfull previous computations.

\subsection{Compare checkpoint files}

The comparison is activated if the option \texttt{-c, --compare} is present in
the command line.

In order to compare two checkpoint files, the markup \texttt{<compare>} has to
be added as a child of the condidered case. In the following exemple, a
checkpoint file comparison is switched on for the case \textit{Grid1}, whereas
no comparision is planed for the case \textit{Grid2}. The comparison is done by
the external script \texttt{cs\_io\_dump} with the option \texttt{--diff}.

\small
\begin{verbatim}
<study label='MyStudy1' status='on'>
    <case label='Grid1' status='on' compute="on" post="off">
        <compare dest="" repo="" threshold="1.0e-5"/>
    </case>
    <case label='Grid2' status='on' compute="off" post="off"/>
</study>
\end{verbatim}
\normalsize

The attributes are:
\begin{list}{$\bullet$}{}

\item \texttt{repo}: id of the results directory in the \textbf{repository} for
example \texttt{repo="20110704-1116"}, if there is a single results directory
in the \texttt{RESU} directory of the case, the id can be ommitted:
\texttt{repo=""};

\item \texttt{dest}: id of the results directory in the \textbf{destination}:
\begin{list}{$\rightarrow$}{}
\item if the id is not known already because the case has not yet run, just let
the attribute empty \texttt{dest=""}, the value will be updated after the run
step in the \textbf{destination} directory (see section \ref{sec:restart});
\item if \textsc{Autovnv} is restarted without the run step (with the command
line \texttt{code\_saturne autovnv -f sample.xml -c} for example), the id of
the results directory in the \textbf{destination} must be given (for example
\texttt{dest="20110706-1523"}), but if there is a single results directory in
the \texttt{RESU} directory of the case, the id can be ommitted:
\texttt{dest=""}, the id will be completed automatically;
\end{list}

\item \texttt{threshold}: real value above which a difference is considered
significant (default: $1e-30$).
\end{list}

Only the attributes \texttt{repo} and \texttt{dest} are mandatory.

\subsection{Run external additional scripts with arguments}

The launch of external scripts is activated if the option \texttt{-p, --post}
is present in the command line.

The markup \texttt{<script>} has to be added as a child of the condidered case.

\small
\begin{verbatim}
<study label='STUDY' status='on'>
    <case label='CASE1' status='on' compute="on" post="on">
        <script label="script_post.py" args="-n 1" dest="" repo="20110216-2147" status="on"/>
    </case>
</study>
\end{verbatim}
\normalsize

The attributes are:
\begin{list}{$\bullet$}{}
\item \texttt{label}: the name of the file of the considered script;
\item \texttt{status}: must be equal to \texttt{on} or \texttt{off}:
activate or desactivate the markup;
\item \texttt{args}: the arguments to pass to the script;
\item \texttt{repo} and \texttt{dest}: id of the results directory in the
\textbf{repository} or in the \textbf{destination};
\begin{list}{$\rightarrow$}{}
\item if the id is not known already because the case has not yet run, just let
the attribute empty \texttt{dest=""}, the value will be updated after the run
step in the \textbf{destination} directory (see section \ref{sec:restart});
\item if there is a single results directory in the \texttt{RESU} directory
(either in the \textbf{repository} or in the \textbf{destination}) of the case,
the id can be ommitted: \texttt{repo=""} or \texttt{dest=""}, the id will be
completed automatically.
\end{list}
If attributes \texttt{repo} and \texttt{dest} exist, their associated value
will be passed to the script as arguments, with options \texttt{"-r"} and
\texttt{"-d"} respectively.
\end{list}

Only the attributes \texttt{label} and \texttt{status} are mandatory.

Several calls of the same script or to different scripts are permitted:
\small
\begin{verbatim}
<study label="STUDY" status="on">
    <case label="CASE1" status="on" compute="on" post="on">
        <script label="script_post.py" args="-n 1" status="on"/>
        <script label="script_post.py" args="-n 2" status="on"/>
        <script label="script_post.py" args="-n 3" status="on"/>
        <script label="another_script.py" status="on"/>
    </case>
</study>
\end{verbatim}
\normalsize

The main objectif of running external scripts is to create or modify
results in order to plot them.

Example of script, which searches printed informations in the listing,
note the function to process the passed command line arguments:
\small
\begin{verbatim}
#!/usr/bin/env python
# -*- coding: utf-8 -*-

import os, sys
import string
from optparse import OptionParser

def process_cmd_line(argv):
    """Processes the passed command line arguments."""
    parser = OptionParser(usage="usage: %prog [options]")

    parser.add_option("-r", "--repo", dest="repo", type="string",
                      help="Directory of the result in the repository")

    parser.add_option("-d", "--dest", dest="dest", type="string",
                      help="Directory of the result in the destination")

    (options, args) = parser.parse_args(argv)
    return options

def main(options):
    m = os.path.join(options.dest, "listing")
    f = open(m)
    lines = f.readlines()
    f.close()

    g = open(os.path.join(options.dest, "water_level.dat"), "w")
    g.write("# time,   h_sim,   h_th\n")
    for l in lines:
       if l.rfind("time, h_sim, h_th") == 0:
           d = l.split()
           g.write("%s  %s  %s\n" % (d[3], d[4], d[5]))
    g.close()

if __name__ == '__main__':
    options = process_cmd_line(sys.argv[1:])
    main(options)
\end{verbatim}
\normalsize

\subsection{Post-processing}

The post-processing is activated if the option \texttt{-p, --post} is present
in the command line.

The following example shows the drawing of four curves (or plots, or 2D lines)
from two files of data (which have the same name \texttt{profile.dat}). There
are two subsets of curves (i.e. frames with axis and 2D lines), in a single
figure. The figure will be saved on the disk in a \textbf{png} format, in the
\texttt{POST} directory of the related study in the \textbf{destination}. Each
drawing of a single curve is defined as a markup child of a file of data inside
a case. Subsets and figures are defined as markup childs of \texttt{<study>}.

\small
\begin{verbatim}
<study label='Study' status='on'>
    <case label='Grid1' status='on' compute="off" post="on">
        <data file="profile.dat" dest="">
            <plot fig="1" xcol="1" ycol="2" legend="Grid level 1" fmt='r-s'/>
            <plot fig="2" xcol="1" ycol="3" legend="Grid level 1" fmt='r-s'/>
        </data>
    </case>
    <case label='Grid2' status='on' compute="off" post="on">
        <data file="profile.dat" dest="">
            <plot fig="1" xcol="1" ycol="2" legend="Grid level 2" fmt='b-p'/>
            <plot fig="2" xcol="1" ycol="3" legend="Grid level 2" fmt='b-p'/>
        </data>
    </case>
    <subplot id="1" legstatus='on' legpos ='0.95 0.95' ylabel="U ($m/s$)" xlabel="Time ($s$)"/>
    <subplot id="2" legstatus='on' legpos ='0.95 0.95' ylabel="U ($m/s$)" xlabel="Time ($s$)"/>
    <figure name="velocity" idlist="1 2" figsize="(4,5)"/>
</study>
\end{verbatim}
\normalsize

\subsubsection{Define curves}

The curves of computational data are build from data files. These data must be
ordered as column and the files should be in results directory in the
\texttt{RESU} directory (either in the \textbf{repository} or in the
\textbf{destination}). Commentaries are allowed in the file, the head of every
commentary line must start with character \texttt{\#}.

In the file of parameters, curves are defined with two markups:
\texttt{<data>} and \texttt{<plot>}:

\begin{list}{$\bullet$}{}
\item \texttt{<data>}: child of markup \texttt{<case>}, defines a file of data;
\begin{list}{$\rightarrow$}{}
\item \texttt{file}: name of the file of data
\item \texttt{repo} or \texttt{dest}: id of the results directory either in the
\textbf{repository} or in the \textbf{destination};
\begin{list}{$\Rightarrow$}{}
\item if the id is not known already because the case has not yet run, just let
the attribute empty \texttt{dest=""}, the value will be updated after the run
step in the \textbf{destination} directory (see section \ref{sec:restart});
\item if there is a single results directory in the \texttt{RESU} directory
(either in the \textbf{repository} or in the \textbf{destination}) of the case,
the id can be ommitted: \texttt{repo=""} or \texttt{dest=""}, the id will be
completed automatically.
\end{list}
\end{list}
The attribute \texttt{file} is mandatory, and either \texttt{repo} or
\texttt{dest} must be present (but not the both) even if it is empty.

\item \texttt{<plot>}: child of markup \texttt{<data>}, defines a single
curve; the attributes are:
\begin{list}{$\rightarrow$}{}
\item \texttt{fig}: id of the subset of curves (i.e. markup \texttt{<subplot>})
where the current curve should be plotted;
\item \texttt{xcol}: number of the column in the file of data for the abscisse;
\item \texttt{ycol}: number of the column in the file of data for the ordinate;
\item \texttt{legend}: add a label to a curve;
\item \texttt{fmt}: format of the line, composed from a symbol, a color and a
linestyle, for example \texttt{fmt="r--"} for a dashed red line;
\item \texttt{xplus}: real to add to all values of the column \texttt{xcol};
\item \texttt{yplus}: real to add to all values of the column \texttt{ycol};
\item \texttt{xfois}: real to multiply to all values of the column
\texttt{xcol};
\item \texttt{yfois}: real to multiply to all values of the column
\texttt{ycol};
\item \texttt{xerr}: draw horizontal error bar (see section \ref{sec:err});
\item \texttt{yerr}: draw vertical error bar (see section \ref{sec:err});
\item some standard options of 2D lines can be added, for example
\texttt{markevery="2"} or \texttt{markersize="3.5"}. These options
are summarized in the table \ref{table:curves}. Note that the options
which are string of characters must be overquoted likes this: 
\texttt{color="'g'"}.

\begin{table}[htbp]
\begin{center}
\begin{tabular}{|l|l|}
\hline
\textbf{Property} & \textbf{Value Type} \\
\hline
alpha & float (0.0 transparent through 1.0 opaque) \\
antialiased or aa & \texttt{True} or \texttt{False} \\
color or c & any matplotlib color \\
dash\_capstyle & \texttt{butt}; \texttt{round}; \texttt{projecting} \\
dash\_joinstyle & \texttt{miter}; \texttt{round}; \texttt{bevel} \\
dashes & sequence of on/off ink in points ex: \texttt{dashes="(5,3)"} \\
label & any string, same as legend\\
linestyle or ls &  \texttt{-}; \texttt{--}; \texttt{-.}; \texttt{:}; \texttt{steps}; ... \\
linewidth or lw & float value in points \\
marker &  \texttt{+}; \texttt{,}; \texttt{.}; \texttt{1}; \texttt{2}; \texttt{3}; \texttt{4}; ... \\
markeredgecolor or mec & any matplotlib color \\
markeredgewidth or mew & float value in points \\
markerfacecolor or mfc & any matplotlib color \\
markersize or ms & float \\
markevery & \texttt{None}; integer; (startind, stride) \\
solid\_capstyle & \texttt{butt}; \texttt{round}; \texttt{projecting} \\
solid\_joinstyle & \texttt{miter}; \texttt{round}; \texttt{bevel} \\
zorder & any number \\
\hline
\end{tabular}
\end{center}
\caption{Options authorized as attributes of the markup \texttt{plot}.}
\label{table:curves}
\end{table}

\end{list}
\end{list}

The attributes \texttt{fig} and \texttt{ycol} are mandatory.

Details on 2D lines properties can be found in the \texttt{matplotlib}
documentation. For more advanced options see section \ref{sec:raw}.

\subsubsection{Define subsets of curves}

A subset of curves is a frame with two axis, axis labels, legend, title and
drawing of curves inside. Such subset is called subplot in the nomenclature
of \texttt{matplotlib}.

\texttt{<subplot>}: child of markup \texttt{<study>}, defines a frame with
severals curves; the attributes are:
\begin{list}{$\rightarrow$}{}
\item \texttt{id}: id of the subplot, should be an integer;
\item \texttt{legstatus}: if \texttt{"on"} display the frame of the legend;
\item \texttt{legpos}: sequence of the relative coordinates of the center of
the legend, it is possible to draw the legend outside the axis;
\item \texttt{title}: set title of the subplot;
\item \texttt{xlabel}: set label for the x axis;
\item \texttt{ylabel}: set label for the y axis;
\item \texttt{xlim}: set range for the x axis;
\item \texttt{ylim}: set range for the y axis.
\end{list}

The attributes \texttt{fig} and \texttt{ycol} are mandatory.

For more advanced options see section \ref{sec:raw}.

\subsubsection{Define figures}

Figure is a compound of subset of curves.

\texttt{<figure>}: child of markup \texttt{<study>}, defines a pictures
with a layout of frames; the attributes are:
\begin{list}{$\rightarrow$}{}
\item \texttt{name}: name of the file to be written on the disk;
\item \texttt{idlist}: list of the subplot to be displayed in the figure;
\item \texttt{title}: add a title on the top of the figure;
\item \texttt{nbrow}: impose a number of row of the layout of the subplots;
\item \texttt{nbcol}: impose a number of column of the layout of the subplots;
\item standard options of figure can be added (table \ref{table:fig}),
for example \texttt{figsize="(3,4)"}.

\begin{table}[htbp]
\begin{center}
\begin{tabular}{|l|l|}
\hline
\textbf{Property} & \textbf{Value Type} \\
\hline
figsize   & width x height in inches; defaults to (4,4) \\
dpi       & resolution; defaults to 200 \\
\hline
\end{tabular}
\end{center}
\caption{Options authorized as attributes of the markup \texttt{figure}.}
\label{table:fig}
\end{table}

Details can be found in the matplotlib documentation.
For more advanced options see section \ref{sec:raw}.

\end{list}

The attributes \texttt{name} and \texttt{idlist} are mandatory.

\subsubsection{Experimental data}

A particular markup is provided for curves of experimental data:
\texttt{<measurement>}; the attributes are:
\begin{list}{$\rightarrow$}{}
\item \texttt{file}: name of the file to be read on the disk;
\item \texttt{path}: path of the directory where the file of experimental data
is. the path could be ommitted (\texttt{path=""}), and in this case, the file
will be searched recursively in the directories of the considered study.
\end{list}

The attributes \texttt{file} and \texttt{path} are mandatory.

In order to draw curves of experimental data, the markup \texttt{<measurement>}
should be used with the markup \texttt{<plot>} as illustrated below:

\small
\begin{verbatim}
<study label='MyStudy' status='on'>
    <measurement file='exp1.dat' path=''>
            <plot fig='1' xcol='1' ycol='2' legend='U Experimental data'/>
            <plot fig='2' xcol='3' ycol='4' legend='V Experimental data'/>
    </measurement>
    <measurement file='exp2.dat' path =''>
            <plot fig='1' xcol='1' ycol='2' legend='U Experimental data'/>
            <plot fig='2' xcol='1' ycol='3' legend='V Experimental data'/>
    </measurement>
    <case label='Grid1' status='on' compute="off" post="on">
        <data file="profile.dat" dest="">
            <plot fig="1" xcol="1" ycol="2" legend="U computed" fmt='r-s'/>
            <plot fig="2" xcol="1" ycol="3" legend="V computed" fmt='b-s'/>
        </data>
    </case>
</study>
<subplot id="1" legstatus='on'  ylabel="U ($m/s$)" xlabel= "$r$ ($m$)" legpos ='0.05 0.1'/>
<subplot id="2" legstatus='off' ylabel="V ($m/s$)" xlabel= "$r$ ($m$)"/>
<figure name="MyFigure" idlist="1 2"  figsize="(4,4)" />
\end{verbatim}
\normalsize

\subsubsection{Curves with error bar}\label{sec:err}

In order to draw horizontal and vertical error bar, it is possible to
specify to the markup \texttt{<plot>} the attributes \texttt{xerr} and
\texttt{yerr} respectively. The value of theses attributes could be:
\begin{list}{$\bullet$}{}
\item a single column number where the average uncertainty is in the file of
data:
\small
\begin{verbatim}
<measurement file='axis.dat' path =''>
    <plot fig='1' xcol='1' ycol='3' legend='Experimental data' xerr='2' />
</measurement>
\end{verbatim}
\normalsize
\item a sequence of two column numbers where the minimum and the maximum
of the uncertainty are in the file of data:
\small
\begin{verbatim}
<data file='profile.dat' dest="">
    <plot fig='1' xcol='1' ycol='2' legend='computation' yerr='3 4' />
</data>
\end{verbatim}
\normalsize
\end{list}

\underline{Notes:}

The attributes \texttt{xerr} and \texttt{yerr} could not be both at the same
time in the markup \texttt{plot}. If horizontal and vertical bars has be
drawn together, repeat the markup \texttt{plot} like this:
\small
\begin{verbatim}
<measurement file='axis.dat' path =''>
    <plot fig='1' xcol='1' ycol='3' legend='Experimental data' xerr='2' />
    <plot fig='1' xcol='1' ycol='3' yerr='4' />
</measurement>
\end{verbatim}
\normalsize

\subsubsection{Matplotlib raw commands}\label{sec:raw}

The file of parameters allows to execute additional matplotlib commands (i.e
Python commands), for curves (2D lines), or subplot, or figure. For every object
drawn, \texttt{Autovnv} associate a name to this object that can be reused in
standard matplotlib commands. Therefore childs markup \texttt{<plt\_command>}
could be added to \texttt{<plot>}, \texttt{<subplot>} or \texttt{<figure>}.

It is possible to add commands with Matlab style or Python style.

\begin{list}{$\bullet$}{}

\item curves or 2D lines: when a curve is drawn, the associated name
are \texttt{line} and \texttt{lines} (with \texttt{line = lines[0]}).

\small
\begin{verbatim}
<plot fig="1" xcol="1" ycol="2" fmt='g^' legend="Simulated water level">
    <plt_command>plt.setp(line, color="blue")</plt_command>
    <plt_command>line.set_alpha(0.5)</plt_command>
</plot>
\end{verbatim}
\normalsize

\item subset of curves: for each subset, the associated name is \texttt{ax}:

\small
\begin{verbatim}
<subplot id="1" legend='Yes' legpos ='0.2 0.95'>
    <plt_command>plt.grid(True)</plt_command>
    <plt_command>plt.xlim(0, 20)</plt_command>
    <plt_command>ax.set_ylim(1, 3)</plt_command>
    <plt_command>plt.xlabel(r"Time ($s$)", fontsize=8)</plt_command>
    <plt_command>ax.set_ylabel(r"Level ($m$)", fontsize=8)</plt_command>
    <plt_command>for l in ax.xaxis.get_ticklabels(): l.set_fontsize(8)</plt_command>
    <plt_command>for l in ax.yaxis.get_ticklabels(): l.set_fontsize(8)</plt_command>
    <plt_command>plt.axis([-0.05, 1.6, 0.0, 0.15])</plt_command>
    <plt_command>plt.xticks([-3, -2, -1, 0, 1])</plt_command>
    <plt_command>plt.subplots_adjust(hspace=0.4, wspace=0.4, right=0.9, left=0.15)</plt_command>
</subplot>
\end{verbatim}
\normalsize

\end{list}

\section{Output and restart}\label{sec:restart}

\textsc{Autovnv} produces several files in the \textbf{destination} directory:
\begin{list}{$\bullet$}{}
\item \texttt{report.txt}: standard output of the script;
\item \texttt{auto\_vnv.log}: log of the code and the \texttt{pdflatex}
compilation;
\item \texttt{report\_global.pdf}: summary of the compilation, run, comparison,
and plot steps;
\item \texttt{report\_detailed.pdf}: details the comparison and display the
plot;
\item \texttt{sample.xml}: udpated file of parameters, useful for restart the
script if an error occurs.
\end{list}

After the computation of a case, if no error occurs, the attribute
\texttt{compute} is set to \texttt{"off"} in the copy of the file of parameters
in the \textbf{destination}. It is allow a restart of \textsc{Autovnv} without
re-run successfull previous computations.
In the same manner, all empty attributes \texttt{repo=""} and \texttt{dest=""}
are completed in the udpated file of parameters.

\section{Tricks}\label{sec:tricks}
\begin{list}{$\bullet$}{}
\item How to comment markups in the file of parameter ?

The opening and closing signs for commantaries are \texttt{<!--} and
\texttt{-->}. In the following example, nothing from the study
\texttt{MyStudy2} will be read:
\small
\begin{verbatim}
<?xml version="1.0"?>
<autovnv>
    <repository>/home/dupond/codesaturne/MyRepository</repository>
    <destination>/home/dupond/codesaturne/MyDestination</destination>

    <study label="MyStudy1" status="on">
        <case label="Grid1" status="on" compute="on" post="on"/>
        <case label="Grid2" status="on" compute="off" post="on"/>
    </study>
    <!--
    <study label="MyStudy2" status="on">
        <case label="k-eps" status="on" compute="on" post="on"/>
        <case label="Rij-eps" status="on" compute="on" post="on"/>
    </study>
    -->
</autovnv>
\end{verbatim}
\normalsize

\item How to add text in a figure ?

It is possible to use raw commands:
\small
\begin{verbatim}
<subplot id='301' ylabel ='Location ($m$)' title='Before jet -0.885' legstatus='off'>
    <plt_command>plt.text(-4.2, 0.113, 'jet')</plt_command>
    <plt_command>plt.text(-4.6, 0.11, r'$\downarrow$', fontsize=15)</plt_command>
</subplot>
 \end{verbatim}
\normalsize

\end{list}


%
\end{document}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
