%-------------------------------------------------------------------------------

% This file is part of Code_Saturne, a general-purpose CFD tool.
%
% Copyright (C) 1998-2017 EDF S.A.
%
% This program is free software; you can redistribute it and/or modify it under
% the terms of the GNU General Public License as published by the Free Software
% Foundation; either version 2 of the License, or (at your option) any later
% version.
%
% This program is distributed in the hope that it will be useful, but WITHOUT
% ANY WARRANTY; without even the implied warranty of MERCHANTABILITY or FITNESS
% FOR A PARTICULAR PURPOSE.  See the GNU General Public License for more
% details.
%
% You should have received a copy of the GNU General Public License along with
% this program; if not, write to the Free Software Foundation, Inc., 51 Franklin
% Street, Fifth Floor, Boston, MA 02110-1301, USA.

%-------------------------------------------------------------------------------

%==================================
%==================================
\section{Keyword list}
%==================================
%==================================
\label{sec:prg_motscles}

The keywords are classified under relevant headings. For each keyword of \CS Kernel,
the following informations are given:

\noindent
\makebox[2.5cm][l]{Variable name}\makebox[1.3cm][l]{Type}%
\makebox[5cm][l]{Allowed values}%
\makebox[4.cm][l]{[Default]}O/C\hspace{1cm}Level\\
\hspace*{2.5cm}Description\\
\hspace*{2.5cm}Potential dependences


\begin{list}{$\bullet$}{}
\item \textbf{Variable name}: Name of the variable containing the keyword.

\item \textbf{Type}: a (Array), i (Integer), r (Real number), c
      (Character string).

\item \textbf{Allowed values}: list or range of allowed values.

\item \textbf{Default}: value defined by the code before any user
      modification (every keyword has one). In some cases, a
      non-allowed value is given (generally $-999$ or $-10^{12}$), forcing the
      user to specify a value. If he does not do it, the code may:
\begin{list}{-}{}
\item automatically use a recommended value (for example, automatic
      choice of the variables for which chronological records will be
      generated).

\item stop, if the keyword is essential.
\end{list}

\item \textbf{O/C}: Optional/Compulsory
\begin{list}{-}{}
\item O: optional keyword, whose default value may be enough.

\item C: keyword which must imperatively be specified.
\end{list}

\item \textbf{Level}: L1, L2 or L3
\begin{list}{-}{}
\item L1 (level 1): the users will have to modify it in the framework of
      standard applications. The L1 keywords are written in bold.

\item L2 (level 2): the users may have to modify it in the framework of
      advanced applications. The L2 keywords are all optional.

\item L3 (level 3): the developers may have to modify it; it keeps its
      default value in any other case. The L3 keywords are all optional.
\end{list}

\item \textbf{Description}:  keyword description, with its potential
      dependences.

\end{list}

The L1 keywords can be modified through the Graphical Use Interface or
in the \texttt{cs\_user\_parameters.f90} file. L2 and L3 keywords can only be modified through
the \texttt{cs\_user\_parameters.f90} file, even if they do not appear in the version proposed
as example it the \texttt{SRC/REFERENCE/base} directory.\\
It is however recommended not to modify the keywords which do not belong to the L1
level.

The alphabetical keyword list is displayed in the index, in the end of
this report.

\minititre{Notes}
$\bullet\ $The notation ``d'' refers to a double precision real. For
           instance, 1.8d-2 means 0.018. \\
$\bullet\ $The notation ``{\tt grand}'' (which can be used in the code)
corresponds to $10^{12}$.

%==================================
\subsection{Input-output}
%==================================

\minititre{Notes}
\begin{list}{$\bullet$}{}
\item Two different files can have neither the same unit number nor the
      same name.
\end{list}

%==================================
\subsubsection{''Calculation'' files}
%==================================

\minititre{General}

%\minititre{1D wall thermal module}
\minititre{Vortex method for LES}

For calculation files related to the vortex method for LES,
please refer to the dedicated \doxygenfile{group__entsor.html}{\texttt{Doxygen} documentation}.

%\minititre{Radiation}
\minititre{Thermochemistry}

For the calculation file related to the thermochemistry, please refer
to the dedicated \doxygenfile{group__userfile.html}{\texttt{Doxygen} documentation}.

%==================================
\subsubsection{Post-processing for {\em EnSight} or other tools}
%==================================

\minititre{Notes}
$\bullet\quad$The format depends on the user choices, and most options
are defined using the GUI or \\
\texttt{cs\_user\_postprocess.c}.\\
$\bullet\quad$The post-processing files can be of the following formats: {\em Ensight Gold},
{\em MED} or {\em CGNS}. The use of the two latter formats depends on
the installation of the corresponding external libraries.\\
$\bullet\quad$For each quantity (problem unknown, preselected numerical
variable or preselected physical parameter), the user specifies if a
post-processing output is wanted. The output frequency can be set.\\

See the dedicated \doxygenanchor{group__entsor.html\#keyvis}{\texttt{Doxygen}
documentation about \texttt{keyvis}}.

%==================================
\subsubsection{Chronological records of the variables on specific points}
%==================================

\minititre{Standard use through Interface or \texttt{cs\_user\_parameters.f90}}
For each quantity (problem unknown, preselected numerical variable or
preselected physical parameter), the user indicates whether chronological records
should be generated, the output period and the position of the
probes. The code generates chronological records at the cell centers located
closest to the geometric points defined by the user by means of their
coordinates. For each quantity, the number of probes and their
index-numbers must be specified (it is not mandatory to generate all
the variables at all the probes).

Please refer to the dedicated \doxygenfile{group__history.html}{\texttt{Doxygen} documentation}.

%==================================
\subsubsection{Time averages}
%==================================
See \doxygenfile{parameters.html}{the dedicated \texttt{Doxygen} documentation}.

%==================================
\subsubsection{Others}
%==================================

For user calculation file, see the following \doxygenfile{group_userfile.html}{\texttt{Doxygen}
documentation}. For other printing options, please refer to the
\doxygenfile{group_entsor.html}{\texttt{Doxygen} documentation} dealing with input/output options.

%==================================
\subsection{Numerical options}
%==================================
\subsubsection{Calculation management}
%==================================

The following \doxygenfile{group__time__step__options.html}{\texttt{Doxygen} documentation}
provides information about the various calculation management options available
in \CS such as \texttt{ntmabs}, \texttt{ntcabs}, etc.

%==================================
\subsubsection{Scalar unknowns}
%==================================

Several keywords refering to the scalar unknowns are
detailed in the following
\doxygenfile{structcs__thermal__model__t.html}{\texttt{Doxygen}
documentation}. The \doxygenfile{structcs__stokes__model__t.html}{\texttt{Doxygen} page}
of the Stokes model structure also contains some keywords such as \texttt{icpsyr},
\texttt{iclvfl} or \texttt{itbrrb}.
For other keywords, please refer to the following \texttt{Doxygen} pages refering to
\doxygenanchor{group__main__variables.html\#nscaus}{\texttt{nscaus}} and
\doxygenanchor{group__scalar__params.html\#iscacp}{\texttt{iscacp}}.

%==================================
\subsubsection{Definition of the equations}
%==================================

For informations about \texttt{istat}, \texttt{iconv}, \texttt{idiff}
or \texttt{idifft}, please refer to the following
\doxygenfile{structcs__var__cal__opt__t.html}{\texttt{Doxygen} documentation}.

Moreover, one can find details about the \texttt{idircl} keyword
\doxygenanchor{group__linear__solver.html\#idircl}{here} and about the
\texttt{ivisse} keyword \doxygenanchor{structcs__stokes__model__t.html\#ivisse}{there}.

%==================================
\subsubsection{Definition of the time advancement}
%==================================

\motcleb{idilat}{i}{1, 2, 3, 4}{1}{O}{L1}
{Algorithm to take into account the density variation in time\\
\hspace*{1.3cm}= 1: steady dilatable flow algorithm (default)\\
\hspace*{1.3cm}= 2: unsteady dilatable flow algorithm\\
\hspace*{1.3cm}= 3: low-Mach number algorithm\\
\hspace*{1.3cm}= 4: non conservative algorithm for fire simulation\\
always useful}

\motcleb{cdtvar}{ra}{strictly positive real number}{1}{O}{L1}
{multiplicative factor applied to the time step for each scalar\\
Hence, the time step used when solving the evolution equation for the
variable is the time step used for the dynamic equations (velocity/pressure)
multiplied by {\tt cdtvar}.\\
The size of the array {\tt cdtvar} is {\tt nvar}. For instance, the multiplicative
coefficient applied to the scalar 2 is {\tt cdtvar(isca(2))}). Yet, the value of
{\tt cdtvar} for the velocity components and the pressure is not used. Also,
although it is possible to change the value of {\tt cdtvar} for the turbulent
variables, it is highly not recommended\\
useful if and only if {\tt nscal} $\geqslant$ 1}

\motcle{varrdt}{r}{strictly positive real number}{0.1}{O}{L3}
{maximum allowed relative increase in the calculated time step value
between two successive time steps (to ensure stability, any decrease in the time step
is immediate and without limit)\\
useful if {\tt idtvar} $\ne$ 0}

For details about time stepping options, please refer to the
dedicated \doxygenfile{group__time__step__options.html}{\texttt{Doxygen}
documentation}.

\minititre{Non-constant time step}
The calculation of the time step uses a reference time step {\tt dtref} (at
the calculation beginning). Later, every time step, the time step value
is calculated by taking into account the different existing limits, in
the following order: \\
\hspace*{1.cm}$\bullet$ {\tt coumax}, {\tt foumax}: the more restrictive limit between
both is used (in the compressible module, the acoustic limitation is added),\\
\hspace*{1.cm}$\bullet$ {\tt varrdt}:  progressive increase and immediate
decrease in the time step,\\
\hspace*{1.cm}$\bullet$ {\tt iptlro}: limitation by the thermal time step,\\
\hspace*{1.cm}$\bullet$ {\tt dtmax} and {\tt dtmin}: clipping of the time step to
the maximum, then to the minimum limit.\\


%==================================
\subsubsection{Turbulence}
%==================================

The $k-\varepsilon$ (standard and linearized production) and $R_{ij}-\varepsilon$
(LRR and SSG) turbulence
models implemented in \CS are ``High-Reynolds'' models. It is therefore
necessary to make sure that the thickness of the first cell neighboring
the wall is larger than the thickness of the viscous sub-layer (at the
wall, $y^+>2.5$ is required as a minimum, and preferably between 30 and
100)\footnote{While creating the mesh, $y^+=\frac{yu*}{\nu}$ is
generally unknown. It can be roughly estimated as $\frac{yU}{10\nu}$, where
$U$ is the characteristic velocity, $\nu$ is the kinematic viscosity of the fluid
 and $y$ is the mid-height of the first cell near the wall.}. If the mesh does
 not respect this condition, the results may be biased
(particularly if thermal processes are involved). Using scalable wall-functions
(cf. keyword {\tt iwallf}) may help avoiding this problem.\\
The v2-f model is a ``Low-Reynolds'' model, it is therefore necessary to
make sure that the thickness of the first cell neighboring the wall is
smaller than the thickness of the viscous sub-layer ($y^+<1$).\\
The $k-\omega$ SST model provides correct results whatever the thickness of the first cell.
Yet, it requires the knowledge of the distance to the wall in every
cell of the calculation domain. The user may refer to the keyword
{\tt icdpar\index{icdpar}} for more details about the potential limitations.\\
The $k-\varepsilon$ model with linear production allows to correct the
known flaw of the standard $k-\varepsilon$ model which overestimates the
turbulence level in case of strong velocity gradients (stopping point).\\
With LES, the wall functions are usually not greatly adapted. It is generally more advisable
(if possible) to refine the mesh towards the wall so that the first cell is in the
viscous sub-layer, where the boundary conditions are simple natural no-slip conditions.\\
Concerning the LES model, the user may refer to the subroutine
\texttt{ussmag} for complements about the dynamic model. Its usage
and the interpretation of its results require particular attention.
In addition, the user must pay further attention when using the dynamic
model with the least squares method based on a partial extended
neighbourhood ({\tt imrgra}=3). Indeed, the results may be degraded if the user
does not implement his own way of averaging the dynamic constant in
\texttt{ussmag} (\textit{i.e.} if the user keeps the local average based
on the extended neighbourhood).\\

For further details, please refer to the following \texttt{Doxygen} documentation
dealing with \doxygenfile{group__turbulence.html}{turbulence options} and
\doxygenfile{group__csttur.html}{turbulence constants}.

%==================================
\subsubsection{Time scheme}
%==================================

By default, the standard time scheme is a first-order.
A second-order scheme is activated automatically with LES modelling.
On the other hand, when ``specific physics'' (gas combustion, pulverised coal,
compressible module) are activated, the second-order scheme is not allowed.

In the current version, the second-order time scheme is not compatible
with the estimators ({\tt iescal}), the velocity-pressure coupling
({\tt ipucou}), the modelling of hydrostatic pressure ({\tt icalhy} and
{\tt iphydr}) and the time- or space-variable time step ({\tt idtvar}).

Also, in the case of a rotation periodicity, a proper second-order is not
ensured for the velocity, but calculations remain possible.

It is recommended to keep the default values of the variables listed
below. Hence, in standard cases, the user does not need to specify these
options.

Please refer to the dedicated
\doxygenfile{group__time__stepping.html}{\texttt{Doxygen} documentation}
for detailed informations about the time stepping parameters.

%==================================
\subsubsection{Gradient reconstruction}
%==================================

The gradient reconstruction keywords such as \texttt{imrgra}, \texttt{nswrgr},
\texttt{epsrgr}, \texttt{imligr}, \texttt{climgr} or \texttt{extrag} are members
of the \texttt{cs\_var\_cal\_opt\_t} structure for which informations can be
found in the following \doxygenfile{structcs__var__cal__opt__t.html}{\texttt{Doxygen}
documentation}.

Details on the \texttt{anomax} keyword can be found
\doxygenanchor{structcs__space__disc__t.html\#anomax}{here} as well.

%==================================
\subsubsection{Solution of the linear systems}
%==================================

See \doxygenfile{parameters.html}{the dedicated \texttt{Doxygen} documentation}
for most settings related to linear solver options.

More informations on these settings can also be found
\doxygenanchor{structcs__var__cal__opt__t.html\#epsilo}{here}.

%==================================
\subsubsection{Convective scheme}
%==================================

For informations on the keywords related to the convective scheme
(i.e. \texttt{blencv}, \texttt{ischcv}, \texttt{isstpc}) please
refer to the following \doxygenfile{structcs__var__cal__opt__t.html}{\texttt{Doxygen}
documentation}.

%==================================
\subsubsection{Pressure-continuity step}
%==================================

Several options related to the pressure-continuity step
are available and can be modified by the user. These options
can be found in the following
\doxygenfile{structcs__stokes__model__t.html}{\texttt{Doxygen} documentation}.
For details about the porosity keyword \texttt{iporos}, please refer to
the dedicated \doxygenanchor{group__additional__source__terms.html\#iporos}{\texttt{Doxygen}
documentation}.

%==================================
\subsubsection{Error estimators for Navier-Stokes}
%==================================

There are currently {\tt nestmx\index{nestmx}}=4 types of local estimators
provided at every time step, with two possible definitions for
each\footnote{Choice made by the user}. These scalars indicate the areas
(cells) in which some error types may be important. They are
stored using the \texttt{cs\_field} API (see
\texttt{field\_get\_val\_s(iestim(iestim), c\_estim)}).
For each estimator, the code writes the minimum and maximum values
in the listing and generates post-processing outputs along with
the other variables.

The additional memory cost is about one real number per cell and per
estimator. The additional calculation cost is variable. For instance, on a
simple test case, the total estimator {\tt iestot} generates an additional cost
of 15 to 20 $\%$ on the CPU time\footnote{Indeed, all the first-order in
space differential terms have to be recalculated at the time $t^{\,n+1}$};
the cost of the three others may be neglected. If the user wants to
avoid the calculation of the estimators during the computation, it is
possible to run a calculation without estimators first, and then activate them on
a restart of one or two time steps.

It is recommended to use the estimators only for visual and qualitative
analysis. Also, their use is compatible neither with a second-order time scheme
nor with a calculation with a frozen velocity field.

{\tt \bf iest = iespre\index{iespre}: prediction} (default name: EsPre).
After the velocity prediction step (yielding $\vect{\widetilde{u}}$), the
estimator $\eta^{\,pred}_{\,i,k}(\vect{\widetilde{u}})$, local variable calculated
at every cell $\Omega_i$, is created from $\vect{\mathcal
R}^{\,pred}(\vect{\widetilde{u}})$, which represents the residual of the equation
solved during this step:
\begin{equation*}
\begin{array}{r c l}
\vect{\mathcal{R}}^{\,pred}(\vect{\widetilde{u}})&= & \rho^n
    \dfrac{\vect{\widetilde{u}}-\vect{u}^n}{\Delta t}
  + \gradt \left(\vect{\widetilde{u}} \right) \cdot \left(\rho  \vect{u}\right)^n
              - \divv \left((\mu+\mu_t)^n \gradt(\vect{\widetilde{u}}) \right)
              + \grad(P^n)     \\
              &- &\text{rest of the right-hand side}
                        (\vect{u}^n, P^n, \text{other variables}^n)
\end{array}
\end{equation*}

By definition:
$$ \eta^{\,pred}_{\,i,k}(\vect{\widetilde{u}})= {|\Omega_i|}^{\,(k-2)/2}\ ||\vect{\mathcal R}^{\,pred}(\vect{\widetilde{u}})||
_{{\mathbb{L}}^{2}(\Omega_i)}$$
%
\begin{itemize}
\item The first family, $k=1$, suppresses the
volume $|\Omega_i|$ which intrinsically appears with the norm
${{\mathbb{L}}^{2}(\Omega_i)}$.
\item The second family, $k=2$, exactly represents the norm
${{\mathbb{L}}^{2}(\Omega_i)}$. The size of the cell therefore
appears in its calculation and induces a weighting effect.
\end{itemize}
$ \eta^{\,pred}_{\,i,k}(\vect{\widetilde{u}})$  is ideally equal to zero when the
reconstruction methods are perfect and the associated system is
solved exactly.

{\tt \bf iest = iesder\index{iesder}: drift}  (default name: EsDer).
The estimator $\eta^{\,der}_{\,i,k}(\vect{u}^{\,n+1})$ is based on the
following quantity (intrinsic to the code):
\begin{equation}
\begin{array}{lll}
 \eta^{der}_{i,k}(\vect{u}^{n+1})
&=& {|\Omega_i|}^{\,(k-2)/2}
||\divs \left(\text{corrected mass flow after the pressure step}\right)
                                              -\ \Gamma||_{{L}^{2}(\Omega_i)} \\
&=& {|\Omega_i|}^{\,(1-k)/2}
|div (\text{corrected mass flow after the pressure step})-\ \Gamma|
\end{array}
\end{equation}
Ideally, it is equal to zero when the Poisson equation related to the pressure is
solved exactly.

{\tt \bf iest = iescor\index{iescor}: correction}  (default name: EsCor).
The estimator $ \eta^{\,corr}_{\,i,k}(\vect{u}^{\,n+1})$ comes directly
from the mass flow calculated with the updated velocity field:
\begin{eqnarray*}
            \eta^{\,corr}_{\,i,k}(\vect{u}^{\,n+1})=
|\Omega_i|^{\,\delta_{\,2,k}}\ |div (\rho^n \vect{u}^{n+1}) -\ \Gamma|
\end{eqnarray*}
The velocities $\vect{u}^{n+1}$ are taken at the cell centers,
the divergence is calculated after projection on the faces.\\
            $ \,\delta_{\,2,k}$ represents the Kronecker symbol.\\
\hspace*{0.5cm}$\bullet$ The first family, $k=1$, is the absolute raw
value of the divergence of the mass flow minus the mass source term.\\
\hspace*{0.5cm}$\bullet$ The second family, $k=2$, represents a physical
property and allows to evaluate the difference in $kg.s^{\,-1}$.\\
Ideally, it is equal to zero when the Poisson equation is solved exactly and
the projection from the mass flux at the faces to the velocity at the cell
centers is made in a set of  functions with null divergence.

{\tt \bf iest = iestot\index{iestot}: total} (default name: EsTot).
The estimator $ \eta^{tot}_{\,i,k}(\vect{u}^{n+1})$, local variable
calculated at every cell $\Omega_i$, is based on the quantity
$\vect{\mathcal R}^{tot}(\vect{u}^{\,n+1})$, which represents the
residual of the equation using the updated values of
$\vect{u}$ and $P$:
\begin{equation*}
\begin{array}{@{}r@{\,} c@{\,} l}
\vect{\mathcal{R}}^{tot}({\vect{u}^{n+1}})&= & \rho^n
    \dfrac{{\vect{u}^{n+1}}-\vect{u}^n}{\Delta t}
  + \gradt \left({\vect{u}^{n+1}} \right) \cdot {\left(\rho  \vect{u}\right)}^{n+1}
              - \divv \left((\mu+\mu_t)^n \gradt({\vect{u}^{n+1}}) \right)
              + \grad({P^{n+1}})     \\
              &- &\text{rest of the right-hand side}
                        ({\vect{u}^{n+1}}, {P^{n+1}}, \text{other variables}^n)
\end{array}
\end{equation*}
%
By definition:
$$ \eta^{tot}_{i,k}(\vect{u}^{n+1})= {|\Omega_i|}^{\,(k-2)/2}\ ||\vect{\mathcal R}^{tot}(\vect{u}^{\,n+1})||
_{\mathbb{L}^{2}(\Omega_i)}$$

The mass flux in the convective term is recalculated from $\vect{u}^{n+1}$
expressed at the cell centres (and not taken from the updated mass flow at the
faces).\\

As for the prediction estimator:
\begin{itemize}
\item The first family, $k=1$, suppresses the
volume $|\Omega_i|$ which intrinsicly appears  with the norm
${\mathbb{L}^{2}(\Omega_i)}$.
\item The second family, $k=2$, exactly represents the norm
${\mathbb{L}^{2}(\Omega_i)}$. The size of the cell therefore
appears in its calculation and induces a weighting effect.
\end{itemize}

The estimators are evaluated depending on the values of {\tt iescal}.

%==================================
\subsubsection{Calculation of the distance to the wall}
%==================================

The options related to the calculation of the distance to the wall are described in
the following \doxygenfile{group__num__wall__distance.html}{\texttt{Doxygen} documentation}.
Some options are used only in the case of the calculation of the non-dimensional distance
to the wall $y^+$ (LES model with van Driest damping). Most of the keywords are simple copies of the
keywords for the numerical options of the general equations, with a potentially
specific value in the case of the calculation of the distance to the wall.\\

%==================================
\subsubsection{Others}
%==================================

Informations concerning the remaining keywords can be reached
through the following \texttt{Doxygen} pages:

\begin{itemize}
\item \doxygenfile{structcs__stokes__model__t.html}{\texttt{iccvfg}
and \texttt{ipucou}}
\item \doxygenfile{structcs__piso__t.html}{\texttt{nterup}
and \texttt{epsup}}
\item \doxygenanchor{structcs__space__disc__t.html\#imvisf}{\texttt{imvisf}}
\item \doxygenfile{structcs__var__cal__opt__t.html}{\texttt{irclu},
\texttt{nswrsm} and \texttt{epsrsm}}
\item \doxygenanchor{group__optcal.html\#isuit1}{\texttt{isuit1}}
\end{itemize}

%====================================================================================
\subsection{Numerical, physical and modelling parameters}
%=============================================================================
\subsubsection{Numeric parameters}
%================================

These parameters correspond to numeric reference values in the code.
They can be used but shall not be modified (they are defined as \texttt{parameter}).

For a list of these physical parameters, please refer to the following
\doxygenfile{group__cstnum.html}{\texttt{Doxygen} documentation}.


%==================================
\subsubsection{Physical parameters}
%==================================

These parameters correspond to physical reference values in the code. They
can be used but shall not be modified (they are defined as {\tt parameter}).

For a list of these physical parameters, please refer to the following
\doxygenfile{group__cstphy.html}{\texttt{Doxygen} documentation}.

%==================================
\subsubsection{Physical variables}
%==================================

Most physical variables are listed in the following
\doxygenfile{group__cstphy.html}{\texttt{Doxygen} documentation}.

Other physical variables such as \texttt{diftl0}, \texttt{srrom},
\texttt{visls0}, \texttt{sigmas} or \texttt{rvarfl} are described in the following
\texttt{Doxygen} pages :
\begin{itemize}
\item \doxygenanchor{group__thermophysical.html\#diftl0}{\texttt{diftl0}},
\item \doxygenanchor{group__enthalpy.html\#srrom}{\texttt{srrom}},
\item \doxygenfile{group__scalar__params.html}{\texttt{visls0}, \texttt{sigmas},
\texttt{rvarfl}}.
\end{itemize}

%==================================
\subsubsection{Modelling parameters}
%==================================

Please refer to the following
\doxygenfile{group__csttur.html}{\texttt{Doxygen} documentation} for
more informations about modelling parameters such as \texttt{xlomlg},
\texttt{almax} or \texttt{uref}.

%==================================
\subsection{ALE}
%==================================

For further details about the ALE calculation options,
please refer to the dedicated \texttt{Doxygen} pages
\doxygenfile{group__albase.html}{here} and
\doxygenfile{group__alstru.html}{there}. The following
\doxygenanchor{group__conv__scheme.html\#iflxmw}{\texttt{Doxygen}
documentation} might be useful as well.

%==================================
\subsection{Thermal radiative transfers: global settings}
%==================================

Most of radiative module keywords may be modified in the user subroutines
\texttt{cs\_user\_radiative\_*} (or, for some of them, through the
thermochemical data files).

For a detailed list of these keywords, please refer to the following
\doxygenfile{structcs__rad__transfer__params__t.html}{\texttt{Doxygen}
documentation}.

%==================================
\subsection{Electric module (Joule effect and electric arcs): specificities}
%==================================

The electric module is composed of a Joule effect module
(\texttt{ippmod(ieljou)\index{ieljou}}) and an electric arcs module
(\texttt{ippmod(ielarc)\index{ielarc}}).

The Joule effect module is designed to take into account the Joule effect
(for instance in glass furnaces) with real or complex potential in the
enthalpy equation. The Laplace forces are not taken into account in the
impulse momentum equation. Specific boundary conditions can be applied to
account for the coupled effect of transformers (offset) in glass furnaces.

The electric arcs module is designed to take into account the Joule effect
(only with real potential) in the enthalpy equation. The Laplace forces
are taken into account in the impulse momentum equation.

The different keywords used in the electric module are detailed in
the following \doxygenfile{structcs__elec__option__t.html}{\texttt{Doxygen}
documentation}.

%=================================================
\subsection{Compressible module: specificities}
%==================================

The keywords used in the global settings are quite few. They are
found in the subroutines \texttt{uscfx1} and \texttt{uscfx2}, in
the \texttt{cs\_user\_parameters.f90} file (see the
description of these user subroutines, \S\ref{prg_uscfx12}).

Detailed informations can be found
\doxygenfile{group__common.html}{here} for the keywords
\texttt{igrdpp}, \texttt{viscv0} and \texttt{icfgrp}.

For \texttt{iviscv}, \texttt{ieos} and \texttt{xmasmr},
please refer to the dedicated
\doxygenfile{structcs__fluid__properties__t.html}{\texttt{Doxygen} documentation}.

%==================================================
\subsection{Lagrangian multiphase flows}
%==================================================
\label{sec:prg_motscles_lagr}

Most of these keywords may be modified in the user subroutines
\texttt{uslag1}, \texttt{uslag2}, \texttt{uslaen},
\texttt{uslast} and \texttt{uslaed}. It is however strongly recommended
not to modify those belonging to the level L3.

First of all, it should be noted that the Lagrangian module is compliant with
almost all the RANS turbulence models and with laminar flows. However, the standard particle
turbulent dispersion model does not take fully advantage of the second-order
$R_{ij}-\varepsilon$ models. The same isotropic model is used as in the
$k-\varepsilon$ models, with $k$ calculated from the trace of $R_{ij}$. Also,
two-way coupling is not compatible with the $k-\omega$ SST model.


\subsubsection{Global settings}

\motcleb{iilagr}{I}{0, 1, 2, 3}{0}{C}{L1}
{activates ($>$0) or deactivates (=0) the Lagrangian module\\
the different values correspond to the following modellings: \\
\hspace*{1.3cm} = 1 Lagrangian two-phase flow in one-way coupling (no influence of
the particles on the continuous phase)\\
\hspace*{1.3cm} = 2 Lagrangian two-phase flow with two-way coupling (influence of
the particles on the dynamics of the continuous phase). Dynamics, temperature and mass may be coupled independently\\
\hspace*{1.3cm} = 3 Lagrangian two-phase flow on frozen continuous phase. This option can
only be used in case of a calculation restart. All the
Eulerian fields are frozen (including the scalar fields). This option
automatically implies \texttt{iccvfg} = 1\\
always useful}

\motcleb{isuila}{i}{0, 1}{0}{C}{L1}
{activation (=1) or not (=0) of a Lagrangian calculation restart.
The calculation restart file read when this option is activated (\texttt{ficaml})
only contains the data related to the particles (see also \texttt{isuist})\\
the global calculation must also be a restart calculation\\
always useful}

\motcleb{isuist}{i}{0, 1}{0}{C}{L1}
{during a Lagrangian calculation restart, indicates whether the particle
statistics (volume and boundary) and two-way coupling terms are to be read from
a restart file (=1) or reinitialised (=0). The file to be read is
\texttt{ficmls}\\
useful if \texttt{isuila} = 1}

\motcleb{nbpmax}{i}{positive or null integer}{1000}{C}{L1}
{maximum number of particles allowed simultaneously in the calculation
domain. It must be reminded that the required memory evolves accordingly}

\motcle{nbpart}{i}{positive or null integer}{0}{O}{L3}
{number of particles treated during one Lagrangian time step\\
\texttt{nbpart} must always be lower than \texttt{nbpmax}\\
always useful, but initialised and updated without intervention of the user}

\motcle{nvls}{i}{integer between 0 and 10}{0}{O}{L2}
{number of additional variables related to the particles\\
the additional variables can be accessed in the arrays
\texttt{eptp\index{eptp}} and \texttt{eptpa\index{eptpa}} by means of the
pointer \texttt{jvls\index{jvls}}: \texttt{eptp(jvls(ii),nbpt)} and
\texttt{eptpa(jvls(ii),nbpt)} (\texttt{nbpt} is
the index-number of the treated particle, and \texttt{ii} an integer
between 1 and \texttt{nvls})}

\motcleb{isttio}{i}{0, 1}{0}{C}{L1}
{indicates the steady (=1) or unsteady (=0) state of the
continuous phase flow\\
in particular, \texttt{isttio} = 1 is needed in order to: \\
\hspace*{1cm}calculate steady statistics in the volume or at the boundaries
(starting respectively from the Lagrangian iterations \texttt{nstist} and
\texttt{nstbor}) \\
\hspace*{1cm}calculate time-averaged two-way coupling source terms (from the
Lagrangian iteration \texttt{nstits}) \\
useful if \texttt{iilagr}=1 or \texttt{iilagr}=2 (if \texttt{iilagr}=3,
then \texttt{isttio}=1 is automatically set)}

\motcleb{injcon}{i}{0, 1}{0}{O}{L1}
{activates (=1) or not (=0) the continuous injection of particles\\
this option allows to inject particles continuously during the duration of the
Lagrangian time step \texttt{dtp\index{dtp}} rather than only once at the
beginning of the Lagrangian iteration. It helps avoiding the fractioning of
the particle cloud close to the injection areas}

\motcleb{iroule}{i}{0, 1}{0}{O}{L1}
{activates (=1) or not (=0) of the particle cloning/fusion technique
(option also called ``Russian roulette'')\\
when \texttt{iroule} = 1, the importance function must be specified {\em via} the
array \texttt{croule\index{croule}} in the user subroutine \texttt{uslaru}}

\motcle{\texttt{ttclag}}{r}{positive real number}{0}{O}{L3}
{physical time of the Lagrangian simulation \\
always useful}

\motcle{iplas}{i}{integer $>$ 0}{1}{O}{L3}
{absolute iteration number (including the restarts) in the Lagrangian
module ({\em i.e.} Lagrangian time step number) \\
always useful}


%==================================================
\subsubsection{Specific physics models associated with the particles}
%==================================================

\motcleb{iphyla}{i}{0, 1, 2}{0}{C}{L1}
{activates ($>$0) or deactivates (=0) the physical models associated to the
particles:\\
\hspace*{1.3cm} = 1: allows to associate with the particles evolution
equations on their temperature (in degrees Celsius), their diameter and
their mass\\
\hspace*{1.3cm} = 2: the particles are pulverised coal particles.
Evolution equations on temperature (in degree Celsius), mass of
reactive coal, mass of char and diameter of the shrinking core are
associated with the particles. This option is available only if the
continuous phase represents a pulverised coal flame\\
always useful}

\motcleb{idpvar}{i}{0, 1}{0}{O}{L1}
{activation (=1) or not (=0) of an evolution equation on the particle
diameter\\
useful if \texttt{iphyla} = 1}

\motcleb{itpvar}{i}{0, 1}{0}{O}{L1}
{activation (=1) or not (=0) of an evolution equation on the particle
temperature (in degrees Celsius)\\
useful if \texttt{iphyla} = 1 and if there is a thermal scalar associated with
the continuous phase}

\motcleb{impvar}{i}{0, 1}{0}{O}{L1}
{activation (=1) or not (=0) of an evolution equation on the
particle mass\\
useful if si \texttt{iphyla} = 1}

\motcleb{tpart}{r}{real number $>$ \texttt{tkelvn}}{700}{O}{L1}
{initialisation temperature (in degree Celsius) for the particles already
present in the calculation domain when an evolution equation on
the particle temperature is activated during a calculation (\texttt{iphyla} =
1 and \texttt{itpvar} = 1)\\
useful if \texttt{isuila} = 1 and \texttt{itpvar} = 0 in the previous
calculation}

\motcleb{cppart}{r}{positive real number}{5200}{O}{L1}
{initialisation value for the specific heat ($J.kg^{-1}.K^{-1}$) of the
particles already present in the calculation domain when an evolution equation
on the particle temperature is activated during a calculation
(\texttt{iphyla} = 1 and \texttt{itpvar} = 1)\\
useful if \texttt{isuila} = 1 and \texttt{itpvar} = 0 in the previous calculation}

\motcleb{iencra}{i}{0, 1}{0}{O}{L1}
{activates (=1) or not (=0) the option of coal particle
fouling. It then is necessary to specify the domain boundaries
on which fouling may take place.\\
useful if \texttt{iphyla} = 2}

\motcleb{tprenc}{r}{real number $>$ \texttt{tkelvn}}{600}{O}{L1}
{limit temperature (in degree Celsius) below which the coal particles do
not cause any fouling (if the fouling model is activated)\\
useful if \texttt{iphyla} = 2 and \texttt{iencra} = 1}

\motcleb{visref}{r}{positive real number}{10000}{O}{L1}
{ash critical viscosity in $kg.m^{-1}.s^{-1}$, in the fouling model
\footnote{J.D. Watt
and T. Fereday (\textit{J.Inst.Fuel}, Vol.42-p99)}\\
useful if \texttt{iphyla} = 2 and \texttt{iencra} = 1}

%==================================================
\subsubsection{Options for two-way coupling}
%==================================================

\motcleb{nstits}{i}{strictly positive integer}{1}{O}{L1}
{number of absolute Lagrangian iterations (including the restarts)
after which a time-average of the two-way coupling source terms is
calculated\\
indeed, if the flow is steady (\texttt{isttio}=1), the average quantities
that appear in the two-way coupling source terms can be calculated over
different time steps, in order to get a better precision\\
if the number of absolute Lagrangian iterations is strictly inferior to
\texttt{nstits}, the code considers that the flow has not yet reached its
steady state (transition period) and the averages appearing in the source
terms are reinitialised at each time step, as it is the case for unsteady
flows (\texttt{isttio}=0)\\
useful if \texttt{iilagr} = 2 and \texttt{isttio} = 1}

\motcleb{ltsdyn}{i}{0, 1}{0}{O}{L1}
{activates (=1) or not (=0) of the two-way coupling on the dynamics of the
continuous phase\\
useful if \texttt{iilagr} = 2 and \texttt{iccvfg} = 0}

\motcleb{ltsmas}{i}{0, 1}{0}{O}{L1}
{activation (=1) or not (=0) of the two-way coupling on the mass\\
useful if \texttt{iilagr} = 2, \texttt{iphyla} = 1 and \texttt{impvar} = 1}

\motcleb{ltsthe}{i}{0, 1}{0}{O}{L1}
{if \texttt{iphyla} = 1 and \texttt{itpvar} = 1, \texttt{ltsthe}
 activates (=1) or not (=0) the two-way coupling on temperature\\
if \texttt{iphyla} = 2, \texttt{ltsthe} activates (=1) or not (=0) the
two-way coupling on the Eulerian variables related to pulverised coal
combustion\\
useful if \texttt{iilagr} = 2}

%==================================================
\subsubsection{Numerical modelling}
%==================================================

\motcle{nordre}{i}{1, 2}{2}{O}{L2}
{order of integration for the stochastic differential equations\\
\hspace*{1.3cm} = 1 integration using a first-order scheme\\
\hspace*{1.3cm} = 2 integration using a second-order scheme\\
always useful}

\motcle{ilapoi}{i}{0, 1}{0}{O}{L3}
{activation (=1) or not (=0) of the solution of a Poisson's equation for
the correction of the particle instantaneous velocities (in order to obtain a
null divergence)\\
this option is not validated and reserved to the development team. The user must not change
the default value}

\motcle{idistu}{i}{0, 1}{1}{O}{L3}
{activation (=1) or not (=0) of the particle turbulent dispersion\\
the turbulent dispersion is compatible only with the RANS turbulent models
($k-\varepsilon$, $R_{ij}-\varepsilon$, v2f or $k-\omega$)\\
(\texttt{iturb}=20, 21, 30, 31, 50 or 60)\\
always useful}

\motcle{idiffl}{i}{0, 1}{0}{O}{L3}
{\texttt{idiffl}=1 suppresses the crossing trajectory effect, making
turbulent dispersion for the particles identical to the turbulent
diffusion of fluid particles\\
useful if \texttt{idistu}=1}

\motcleb{modcpl}{i}{positive integer}{0}{O}{L1}
{activates ($>$0) or not (=0) the complete turbulent dispersion model\\
when \texttt{modcpl} is strictly positive, its value is interpreted as the
absolute Lagrangian time step number (including restarts) after which the
complete model is applied\\
since the complete model uses volume statistics, \texttt{modcpl} must
either be 0 or be greater than \texttt{idstnt}\\
useful if \texttt{istala} = 1}

\motcleb{idirla}{i}{1, 2, 3}{1}{O}{L1}
{$x$, $y$ or $z$ direction of the complete model\\
it corresponds to the main directions of the flow\\
useful if \texttt{modcpl} $>$ 0}

%==================================================
\subsubsection{Volume statistics}
%==================================================

\motcleb{istala}{i}{0, 1}{0}{C}{L1}
{activation (=1) or not (=0) of the calculation of the volume
statistics related to the dispersed phase\\
if \texttt{istala} = 1, the calculation of the statistics is activated
starting from the absolute iteration (including the restarts) \texttt{idstnt}\\
by default, the statistics are not steady (reset to zero at every
Lagrangian iteration). But if \texttt{isttio}=1, since the flow is steady,
the statistics will be averaged overt he different time steps\\
the statistics represent the significant results on the particle cloud\\
always useful}

\motcleb{seuil}{r}{positive real number}{0}{O}{L1}
{every cell of the calculation domain contains a certain quantity of
particles, representing a certain statistical weight (sum of the
statistical weights of all the particles present in the cell). \texttt{seuil}
is the limit statistical weight value, below which the contribution of the
cell in term of statistical weight is not taken into account in the volume
statistics (for the complete turbulent dispersion model, in the
Poisson's equation used to correct the mean velocities or in the listing and
post-processing outputs)\\
useful if \texttt{istala} = 1}

\motcleb{idstnt}{i}{strictly positive integer}{1}{C}{L1}
{absolute Lagrangian iteration number (including the restarts) after
which the calculation of the volume statistics is activated\\
useful if \texttt{istala} = 1}

\motcleb{nstist}{i}{integer $\geqslant$ \texttt{idstnt}}{\texttt{idstnt}}{O}{L1}
{absolute Lagrangian iteration number (including the restarts) after
which the volume statistics are cumulated over time (they are then said to be
steady)\\
if the absolute Lagrangian iteration number is less than \texttt{nstist},
or if the flow is unsteady (\texttt{isttio}=0), the statistics are reset
to zero at each Lagrangian iteration (the volume statistics are then said
to be unsteady)\\
useful if \texttt{istala}=1 and \texttt{isttio}=1}

\motcleb{nomlag}{ca}{string of less than 50 characters}{VarLagXXXX}{O}{L1}
{name of the volumetric statistics, displayed in the listing
and the post-processing files. The default value is given above, with ``XXXX''
representing a four digit number (for instance 0001, 0011 ...)\\
useful if \texttt{istala} = 1\\
{\em Warning: this name is also used to reference information in the restart file
\mbox{(\texttt{isuist} =1)}. If the name of a variable is changed between two
calculations, then it will not be possible to read its value from the restart file}}

\motcleb{nvlsts}{i}{0 $\leqslant$ integer $\leqslant$ \texttt{nussta}=20}{0}{O}{L1}
{number of additional user volume statistics\\
the additional statistics (or their cumulated value in the steady
case) can be accessed in the array \texttt{statis} by means of the pointer
\texttt{ilvu\index{ilvu}}: \texttt{statis(iel,ilvu(ii)\index{statis}})
(\texttt{iel} is the cell index-number and \texttt{ii} an integer between
1 and \texttt{nvlsts})\\
useful if \texttt{istala} = 1}

\motcle{npst}{i}{positive integer}{0}{O}{L3}
{number of iterations during which steady volume statistics have
been cumulated\\
useful if \texttt{istala}=1, \texttt{isttio}=1 and if \texttt{nstist} is
less than or equal to the current Lagrangian iteration\\
\texttt{npst} is initialised and updated automatically by the code, its
value is not to be modified by the user}

\motcle{npstt}{i}{positive integer}{0}{O}{L3}
{number of iterations during which volume statistics have been
calculated (the potential iterations during which unsteady
statistics have been calculated are counted in \texttt{npstt})\\
useful if \texttt{istala}=1\\
\texttt{npstt} is initialised and updated automatically by the code,
its value is not to be modified by the user}

\motcle{tstat}{r}{positive real number}{\texttt{dtp}}{O}{L3}
{if the volume statistics are calculated in a steady way, \texttt{tstat}
represents the physical time during which the statistics have been cumulated\\
if the volume statistics are calculated in a unsteady way,
then \texttt{tstat=dtp\index{dtp}} (it is the Lagrangian time step, because the
statistics are reset to zero at every iteration)\\
useful if \texttt{istala}=1\\
\texttt{tstat} is initialised and updated automatically by the code,
its value is not to be modified by the user}

%==================================================
\subsubsection{Display of particles and trajectories}
%==================================================

The definition of the particle visualization output meshes
itself is done with that of all other postprocessing mesh zones, either
using the GUI, or the \texttt{cs\_user\_postprocess\_meshes} function of
\texttt{cs\_user\_postprocess.c}.

The following options determine which variables are output on those particle
or particle trajectory segment output meshes for which the ``automatic output''
is activated.

\motcleb{ivisv1}{i}{0, 1}{0}{O}{L1}
{associates (=1) or not (=0) the variable ``velocity of the locally
undisturbed fluid flow field'' with the output of particles or trajectories\\
always useful}

\motcleb{ivisv2}{i}{0, 1}{0}{O}{L1}
{associates (=1) or not (=0) the variable ``particle velocity''
with the output of particles or trajectories\\
always useful}

\motcleb{ivistp}{i}{0, 1}{0}{O}{L1}
{associates (=1) or not (=0) the variable ``residence time''
with the output of particles or trajectories\\
always useful}

\motcleb{ivisdm}{i}{0, 1}{0}{O}{L1}
{associates (=1) or not (=0) the variable ``particle diameter''
with the output of particles or trajectories\\
always useful}

\motcleb{iviste}{i}{0, 1}{0}{O}{L1}
{associates (=1) or not (=0) the variable ``particle temperature''
with the output of particles or trajectories\\
always useful}

\motcleb{ivismp}{i}{0, 1}{0}{O}{L1}
{associates (=1) or not (=0) the variable ``particle mass''
with the output of particles or trajectories\\
always useful}

\motcleb{ivisdk}{i}{0, 1}{0}{O}{L1}
{associates (=1) or not (=0) the variable ``shrinking core diameter of
the coal particles'' with the output of particles or trajectories\\
useful only if \texttt{iphyla} = 2}

\motcleb{ivisch}{i}{0, 1}{0}{O}{L1}
{associates (=1) or not (=0) the variable ``mass of reactive coal of the
coal particles'' with the output of particles or trajectories\\
useful only if \texttt{iphyla} = 2}

\motcleb{ivisck}{i}{0, 1}{0}{O}{L1}
{associates (=1) or not (=0) the variable ``mass of coal of the
coal particles'' with the output of particles or trajectories\\
useful only if \texttt{iphyla} = 2}

%==================================================
\subsubsection{Display of the particle/boundary interactions and the statistics at the boundaries}
%==================================================

\motcleb{iensi3}{i}{0, 1}{0}{C}{L1}
{activates (=1) or not (=0) of the recording of the particle/boundary
interactions in  \texttt{parbor\index{parbor}}, and of the calculation of the
statistics at the corresponding boundaries, for post-processing
(\textit{EnSight6} format)\\
By default, the statistics are unsteady (reset to zero at every
Lagrangian iteration). They may be steady if \texttt{isttio}=1 ({\em i.e.}
calculation of a cumulated value over time, and then calculation of an
average over time or over the number of interactions with the boundary)\\
always useful}

\motcleb{nstbor}{i}{strictly positive integer}{1}{O}{L1}
{number of absolute Lagrangian iterations (including the restarts)
after which the statistics at the boundaries are considered steady and are
averaged (over time or over the number of interactions)\\
If the number of absolute Lagrangian iterations is less than \texttt{nstbor},
or if \texttt{isttio}=0, the statistics are reset to zero at every
Lagrangian iteration (unsteady statistics)\\
useful if \texttt{iensi3}=1 and \texttt{isttio}=1}

\motcleb{seuilf}{r}{positive real number}{0}{O}{L1}
{every boundary face of the mesh undergoes a certain number of
interactions with particles, expressed in term of statistical weight
(sum of the statistical weights of all the particles which have
interacted with the boundary face). \texttt{seuilf} is
the limit statistical weight value, below which the contribution of the
face is not taken into account in the
statistics at the boundaries for post-processing\\
useful if \texttt{iensi3}=1}

\motcleb{inbrbd}{i}{0, 1}{1}{O}{L1}
{activates (=1) or not (=0) of the recording of the number of particle/boundary
interactions, and of the calculation of the associated boundary statistics.\\
\texttt{inbrd} = 1 is a compulsory condition to use the particulate average
\texttt{imoybr} = 2\\
useful if \texttt{iensi3}=1}

\motcleb{iflmbd}{i}{0, 1}{0}{O}{L1}
{activates (=1) or not (=0) of the recording of the particulate mass flow
related to the particle/boundary interactions, and of the calculation of
the associated boundary statistics\\
\texttt{inbrd} = 1 is a compulsory condition to use \texttt{iflmbd}=1\\
useful if \texttt{iensi3}=1 and \texttt{inbrbd}=1}


\motcleb{iangbd}{i}{0, 1}{0}{O}{L1}
{activates (=1) or not (=0) of the recording of the angle between a
particle trajectory and a boundary face involved in a particle/boundary
interaction, and of the calculation of the associated boundary statistics\\
useful if \texttt{iensi3}=1}

\motcleb{ivitbd}{i}{0, 1}{0}{O}{L1}
{activates (=1) or not (=0) of the recording of the velocity of a particle
involved in a particle/boundary interaction, and of the calculation of
the associated boundary statistics\\
useful if \texttt{iensi3}=1}

\motcleb{iencbd}{i}{0, 1}{0}{O}{L1}
{activates (=1) or not (=0) of the recording of the mass of coal particles
stuck to the wall due to fouling, on the boundary faces of the \texttt{iencrl}
interaction type\\
useful if \texttt{iensi3}=1, \texttt{iphyla}=2, \texttt{iencra}=1, and if
there is at least one boundary face of the \texttt{iencrl} interaction
type}

\motcleb{nusbor}{i}{positive integer}{0}{O}{L1}
{number of additional user data to record for the calculation
of additional boundary statistics in \texttt{parbor\index{parbor}}\\
useful if \texttt{iensi3}=1}

\motcleb{nombrd}{ca}{string of less than 50 characters}{see
\texttt{uslag1}}{O}{L1}
{name of the boundary statistics, displayed in the listing
and the post-processing files\\
useful if \texttt{iensi3}=1\\
{\em Warning: this name is also used to reference information in the restart file
\mbox{(\texttt{isuist} =1)}. If the name of a variable is changed between two
calculations, it will not be possible to read its value from the restart file}}

\motcleb{imoybr}{ia}{0, 1, 2}{0 , 1 or 2}{O}{L1}
{the recordings in \texttt{parbor} at every particle/boundary interaction are
cumulated values (possibly reset to zero at every iteration in the
unsteady case). They must therefore be divided by a quantity to
get boundary statistics. The user can choose between two average types:\\
\hspace*{1.3cm} = 0: no average is applied to the recorded cumulated values\\
\hspace*{1.3cm} = 1: a time-average is calculated. The cumulated value
is divided by the physical duration in the case of steady
averages (\texttt{isttio}=1). The cumulated value is divided by the value of
the last time step in the case of unsteady averages (\texttt{isttio}=0),
and also in the case of steady averages while the
absolute Lagrangian iteration number is inferior to \texttt{nstbor}\\
\hspace*{1.3cm} = 2: a particulate average is calculated. The cumulated
value is divided by the number of particle/boundary interactions (in terms of
statistical weight) recorded in \texttt{parbor(nfabor,inbr)}. This average
can only be calculated when \texttt{inbrbd}=1. The average is calculated if
the number of interactions (in statistical weight) of the considered
boundary face is strictly higher than \texttt{seuilf}, otherwise the average
at the face is set to zero\\
only the cumulated value is recorded in the restart file\\
useful if \texttt{iensi3}=1}

\motcle{npstf}{i}{positive integer}{0}{O}{L3}
{number of iterations during which steady boundary statistics have
been cumulated\\
useful if \texttt{iensi3}=1, \texttt{isttio}=1 and \texttt{nstbor} less than
or equal to the current Lagrangian iteration\\
\texttt{npstf} is initialised and updated automatically by the code,
its value is not to be modified by the user}

\motcle{npstft}{i}{positive integer}{0}{O}{L3}
{number of iterations during which boundary statistics have
been calculated
(the potential iterations during which unsteady
statistics have been calculated are counted in \texttt{npstft})\\
useful if \texttt{iensi3}=1\\
\texttt{npstft} is initialised and updated automatically by the code,
its value is not to be modified by the user}

\motcle{tstatp}{r}{positive real number}{\texttt{dtp}}{O}{L3}
{if the recording of the boundary statistics is steady, \texttt{tstatp}
contains the cumulated physical duration of the recording of the boundary
statistics\\
if the recording of the boundary statistics is unsteady, then
\texttt{tstat=dtp\index{dtp}} (it is the Lagrangian time step, because the
statistics are reset to zero at every time step)\\
useful if \texttt{iensi3}=1}
