%-----------------------------------------------------------------------
%
%     This file is part of the Code_Saturne Kernel, element of the
%     Code_Saturne CFD tool.
%
%     Copyright (C) 1998-2008 EDF S.A., France
%
%     contact: saturne-support@edf.fr
%
%     The Code_Saturne Kernel is free software; you can redistribute it
%     and/or modify it under the terms of the GNU General Public License
%     as published by the Free Software Foundation; either version 2 of
%     the License, or (at your option) any later version.
%
%     The Code_Saturne Kernel is distributed in the hope that it will be
%     useful, but WITHOUT ANY WARRANTY; without even the implied warranty
%     of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
%     GNU General Public License for more details.
%
%     You should have received a copy of the GNU General Public License
%     along with the Code_Saturne Kernel; if not, write to the
%     Free Software Foundation, Inc.,
%     51 Franklin St, Fifth Floor,
%     Boston, MA  02110-1301  USA
%
%-----------------------------------------------------------------------
%
\programme{co**, cp**, fu** and so ... d3p*, ebu*, lwc*, pdf*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Fonction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

From a CFD point of vue combustion is a (somestimes very) complicated
way to determine $\rho$.  Models needs few extra fields of scalar with
regular transport equation, somes of them with a reactive or
interfacial source term.

Modeling of combustion is able to deal with gaz phase combustion
(diffusion, premix, partial premix), and with solid or liquid fuels.

Combustion of condensed fuels involves one-way interfacial flux due to
phenomenon in the condensed phase (evaporation or pyrolisis) and
reciprocal ones (heterogeneous combustion). Many of the species
injected in the gaz phase are afterwards involved in gaz phase
combsution.

That is the reason why a lot of modules are similar for gaz, coal and
fuel combusiton modelling. Obviously, the thermodynamical description
of gaz species is similar in every version as close as possible of the
JANAF rules.

Every models are developped in both an adiabatic version and an
undiabatic (permeatic) one, so in addition with standard, the rule to
call models is :
IPPMOd(index model) = -1   unused
IPPMOD(index model) =  0   simplest adiabatic version
IPPMOD(index model) =  1   simplest permeatic version
Eventually
IPPMOD(index model) = 2.p   p° adiabatic version
IPPMOD(index model) = 2.p+1  P° permeatic version

Every permeatic version involves the transport of enthalpy (one more variable).

%=================================
\subsection{Gaz combustion modelling}
%=================================

Combustion of gaz is limited by disponibility (in the same fluid
particle) of both fuel and oxidant and by kinetic effects (a lot of
chemical reactions involved can be described using an Arrhenius law
with an high activation energy). The mixing of mass (atoms) incoming
with fuel and oxydant is described by a mixture fraction (mass
fraction of mass incoming with fuel), this variable is not affected by
combustion. A progress variabl is used to describe the transformation
of the mixture from fuel and oxydant to product (carbon dioxyde and so
on).

Combsution of gaz is, traditionnaly, splitted in premix and diffusion.
In premix combustion process a first stage of mixing have been
realised (without blast ...) and the mixture is introduced in the
boiler (or combustion can). In industrial common conditions the
combustion is mainly limited by the mixing of fresh gaz (inert) and
burnt ones resulting in the inflammation of the first and their
conversion to burnt ones ; so an assumption of chemistry much faster
than mixing induces an intermittent regime. The gaz flow is
constituted of totally fresh and totally burnt gaz (the flamme
containing the gaz during their transformation) is "extremuly"
%thin. With previous assumptions, Spalding \ref{Comb} have established
thin. With previous assumptions, Spalding have established
the "Eddy Break Up" model, wich allows a complete descrition with only
one progress variable (mixture fraction is homogeneous).  In diffusion
flame the fuel and the oxydant are introduced by two (at least)
inlets, in common industiral conditions, their mixing is the main
limitation and the mixture fraction is enough to qualify a fluid
particle, but in turbulent flow a probability density function of the
mixture fraction is needed to qualify the thermodynamical state of the
bulk. So both the mean and the variance of the mixture fraction are
needed (two variables).

In the real world the chemistry is not so fast and, often, the mixing
is not as homogeneous as wished. Then the industrial combustion occurs
in partial premix combsution. Partial premix occurs if the mixing is
not finished (at molecular level) when the mixing is introduced, or if
air, or fuel, are staggered, or if a diffusion flame is blowed
off. For these situations, and specifically for lean premix gaz
%turbines Libby \& Williams \ref{Comb} have developped a model allowing
turbines Libby \& Williams have developped a model allowing
a descirption of both mixing and chemical limitations. A collaboration
%between the LCD Poitiers \ref {Comb} and EDF R\&D allows a simpler
between the LCD Poitiers and EDF R\&D allows a simpler
version of their algorithm. Not only the mean and the variance of both
mixture fraction and progress variable are needed but so the
covariance (five variables).


%=================================
\subsection{Coal combustion modelling}
%=================================

Combustion of coal is the main way to produce electricity in the
world.  The coal is a natural product with a very complex composition,
during the industrial process of milling the raw coal is broken in
tiny particles of different sizes. After its introduction in the
boiler, the caol particles undergoes drying, devolatilisation (heating
of coal turn it in a mixture of char and gases), heterogenous
combustion (of char leaving to carbone monoxide) leaving an ash
particle. Each of therse phenomena are taken in account for some class
of particles : a class is caracterised by a coal (it is useful to burn
mixture of coals with differents ranks or mixture of coal with
biomasse ...) and an initla diameter. For each class, \CS computes the
number and the mass of particles by unit mass of mixture.  The main
assumption is to solve only one speed (and pressure) field : it means
the discrepancy of speed between coal particles and gases is supposed
negligible.  Due to the radiation and heterogeneous combustion,
temperature can be different for gas and different size particles : so
the specific enthalpy of each particle class is solved.  The
%description of coal pyrolisis proposed by Kobayashi \cite{Comb} and
%bhayakar \cite{Comb} is used, leaving at two source terms for light
descripton of coal pyrolisis proposed by Kobayashi and
bhayakar is used, leaving at two source terms for light
and heavy volatile matters (the moderate temperature reaction produces
gases with low molecular mass, the high temperature reaction produces
heavier gases and less char) represented by passive scalar : mixture
fraction.  The description of heterogeneous reaction (who produce
carbon monoxide) leads to a source term for the carbon : a mixture
fraction who can't be greater than the results of stoechiometric
oxydation of char by air.  The retained model for the gaz phase
combusiton is diffusion flammelets surrounding each particle, so the
previous gaseous fuels are mixed in a local mean fuel and the mixing
with air is represented by a pdf between air and mean local fuel
constructed with the variance of a passive scalar linked with air
(interfacial mass flux produce a source term for this scalar).




%=================================
\subsection{Heavy Fuel Oil combustion modelling}
%=================================

Combustion of heavy fuel oil have been hugely used to produced
electrical energy. Environemental regulation turns it more difficult
and less acceptable, a focus is needed on pollutant emission mainly
soulphur oxide and particles (condensation of sulphuric acid can
aggregate soot).  The description of fuel evaporation is done with
respect of its heaviness : after a minimum temperature is reached, the
gain of enthalpy is distributed between heating and evaporation. This
way the evaporation takes place on a range of temperature (which can
be large). The "total" evaporation is common for light oil but
impossible for heavy ones, so a particle similar to char is leaved ;
the heterogeneous oxydation of this char particle is very similar to
coal char ones.  Injection of fuel is described (2006 version) with
only one class of particle, the number and mass of particles is
calculated eveywhere. And so the enthalpy. So three variables are used
to describe the condensed pahse. Like for coal, only one speed field
is solved.  The model for gas combustion is very similar to coal ones
but a special is paid to sulphur (assumed to leave the particle as H2S
during evaporation and to be converted to SO2 during gas combustion).


%==================================
%==================================
%\section{Bibliography}
%==================================
%==================================
%\begin{thebibliography}{3}

%\bibitem{Comb} {\sc Plion P., {\em et al.}},\\
%{\em Le titre},\\
%Les référence.


%\end{thebibliography}
